AbxD01 Analysis

### 2015-01-25 abxD01_analysis.Rmd
fixed figure 2

This figure was built using the R script called "abxD01.barcharts.xOTU.test.r" in the abxD01 folder.The input files are listed in Github. The graph can be sorted 2 ways either by the untreated's decreasing relabund, or by the most abundant within each phylum. For each treatment, the resulting *C. difficile* CFU/g Feces is also labeled on the Y axis below the treatment name. File saved as "topdose2_tx1_sorted_10x20.pdf" and "topdose2_tx2_sorted_10x20.pdf"."

@@

meta <- read.table('~/Desktop/mothur/abxD01/correlation/abxD01.final.an.unique_list.0.03.subsample.filter16mintotal.shared.correl.topdose2.txt',header=T)
#meta<-meta[1:96] #change based on number of OTUs above .05%, then add 2 for first two cols
c<-1
otu <- c()
cor.spear <- c()
pval.spear <- c()
cor.ken = c()
pval.ken = c()
for(i in 3:length(meta)){
  otu[c] <- colnames(meta[i])
  cor.spear[c] <- cor.test(meta[,2],meta[,i], method="spearman")$estimate
  pval.spear[c] <- cor.test(meta[,2],meta[,i], method="spearman")$p.value
  cor.ken[c] <- cor.test(meta[,2],meta[,i], method="kendall")$estimate #good to see because kendall handles ties
  pval.ken[c] <- cor.test(meta[,2],meta[,i], method="kendall")$p.value #but only works if this is tao-b and not tao-a which im not sure about
  c <- c+1
}

@@

results = NULL
results <- matrix(c(otu, cor.spear, pval.spear, cor.ken, pval.ken), ncol=5)
colnames(results) <- c('otu','corSpear','pvalSpear', "corKen", "pvalKen")
results <- results[order(results[,3]),]  #order by pvalue column=3
write.table(results[1:dim(results)[1],], file="~/Desktop/mothur/abxD01/correlation/abxD01.final.an.unique_list.0.03.subsample.filter16mintotal.shared.correl.topdose2.results.txt", sep="\t", row.names=FALSE)

@@

##Paper Figure: correlations for topdose 
c<-read.csv("~/Desktop/mothur/abxD01/correlation/corr_allSig_topdose2_stripchart.csv", header=T)
labels=c("Lachnospiraceae", "Ruminococcaceae", "Clostridia", "Lactobacillales", "Firmicutes", "Bacillales", "Porphyromonadaceae", "Bacteroidales", "Bacteroidetes", "Actinobacteria", "Proteobacteria", "Anaeroplasma", "Deinococcus", "Unclassified")
ns=c("n=18", "n=9", "n=8", "n=4", "n=2", "n=1", "n=15", "n=4", "n=1", "n=4", "n=3", "n=1", "n=1", "n=3")

par(mar=c(12, 7, 2, 2) +0.1, mgp=c(5, 1, 0)) #default is 5.1 4.1 4.1 2.1 [bottom, left, top, right space], mgp=c(3, 1, 0) [label line location for x/y location labels, tick mark labels location, tick mark locations]
stripchart(c$cor~c$order, vertical=TRUE, ylab="Correlation", ylim=c(.85, -.85), cex.axis=1.1, cex=2, pch=21, lwd=3, col="black", xaxt="n", method="jitter",  cex.lab=1.7, yaxt="n")
axis(1, cex.axis=1, at=(1:nlevels(c$name)), labels=FALSE)
text(1:nlevels(c$name)+.1, par("usr")[3]-0.16, label=labels, xpd=NA, pos=2, srt=45, cex=1.2)
text(1:nlevels(c$name), par("usr")[3]-0.01, label=ns, xpd=NA, pos=1, cex=.9)
axis(2, cex.axis=1.1, at=seq(-0.85,+0.85, by=.1), las=1)
abline(v=c(1:nlevels(c$name)), col="dark gray")
#abline(h=c(.3, -.3), col="dark gray", lty="dashed", lwd=4) #change to something else
abline(h=0, col="black", lwd=2)

@@

Figure saved as "corr_allSig_topdose2_10x8.pdf".

@@

The code for this is still in the works. But is uploaded to github under Figure 3 folder.

Input files on github:
1. abxD01.final.tx.2.subsample.allvanctitr.forlogscale.csv
2. abxD01/barcharts/allvanctitr_tx2_barchart_ids.csv

>>>>

This figure was built using the R script called "abxD01.barcharts.xOTU.test.r" in the abxD01 folder.The input files are listed in Github. The graph can be sorted 2 ways either by the untreated's decreasing relabund, or by the most abundant within each phylum. For each treatment, the resulting *C. difficile* CFU/g Feces is also labeled on the Y axis below the treatment name. 
The parameters I changed in the r script were:
sortbyphyl<-TRUE
graphbyphyl<-FALSE
abx<-c("Untreated", "Ciprofloxacin", "Clindamycin", "Vancomycin", "Streptomycin", "Cefoperazone", "Ampicillin", "Metronidazole")

File saved as "topdose2_tx1_sorted_10x20.pdf" and "topdose2_tx2_sorted_10x20.pdf"."

@@

meta <- read.table('~/Documents/Github/abxD01/Figure 2/abxD01.final.an.unique_list.0.03.subsample.filter16mintotal.shared.correl.topdose2.txt',header=T)
#meta<-meta[1:96] #change based on number of OTUs above .05%, then add 2 for first two cols
c<-1
otu <- c()
cor.spear <- c()
pval.spear <- c()
cor.ken = c()
pval.ken = c()
for(i in 3:length(meta)){
  otu[c] <- colnames(meta[i])
  cor.spear[c] <- cor.test(meta[,2],meta[,i], method="spearman")$estimate
  pval.spear[c] <- cor.test(meta[,2],meta[,i], method="spearman")$p.value
  cor.ken[c] <- cor.test(meta[,2],meta[,i], method="kendall")$estimate #good to see because kendall handles ties
  pval.ken[c] <- cor.test(meta[,2],meta[,i], method="kendall")$p.value #but only works if this is tao-b and not tao-a which im not sure about
  c <- c+1
}

@@

results = NULL
results <- matrix(c(otu, cor.spear, pval.spear, cor.ken, pval.ken), ncol=5)
colnames(results) <- c('otu','corSpear','pvalSpear', "corKen", "pvalKen")
results <- results[order(results[,3]),]  #order by pvalue column=3
write.table(results[1:dim(results)[1],], file="~/Documents/Github/abxD01/Figure 2/abxD01.final.an.unique_list.0.03.subsample.filter16mintotal.shared.correl.topdose2.results.txt", sep="\t", row.names=FALSE)

@@

##Paper Figure: correlations for topdose 
otuCorrs<-read.csv("~/Documents/Github/abxD01/Figure 2/corr_avgAbove0.001_topdose2_stripchart.csv", header=T)
#labels=c("Lachnospiraceae", "Ruminococcaceae", "Clostridia", "Lactobacillales", "Firmicutes", "Bacillales", "Porphyromonadaceae", "Bacteroidales", "Bacteroidetes", "Actinobacteria", "Proteobacteria", "Anaeroplasma", "Deinococcus", "Unclassified")
#ns=c("n=18", "n=9", "n=8", "n=4", "n=2", "n=1", "n=15", "n=4", "n=1", "n=4", "n=3", "n=1", "n=1", "n=3")

#test<-otuCorrs[otuCorrs$phyl=="Bacteroidetes",]
#length(unique(test$family))

pch <- c(1,1,1,1, 6,6, 6, 6, 6, 6, 6, 6,0, 0, 9, 2, 2, 4)
par(mar=c(12, 7, 2, 2) +0.1, mgp=c(4, 1, 0)) #default is 5.1 4.1 4.1 2.1 [bottom, left, top, right space], mgp=c(3, 1, 0) [label line location for x/y location labels, tick mark labels location, tick mark locations]
stripchart(otuCorrs$corr ~ otuCorrs$graphOrder, vertical=TRUE, ylab="Spearman Correlation", ylim=c(-1, 1), cex.axis=1.1, cex=2, pch=pch, lwd=2, col="black",  method="jitter",  jitter=0.25, cex.lab=1.3, yaxt="n", xaxt="n")
axis(1, cex.axis=1, at=(1:nlevels(otuCorrs$family)), labels=FALSE)
text(1:nlevels(otuCorrs$family)+.1, par("usr")[3]-0.05, label=unique(otuCorrs$family), xpd=NA, pos=2, srt=45, cex=1.2)
#text(1:nlevels(otuCorrs$family), par("usr")[3]-0.01, label=ns, xpd=NA, pos=1, cex=.9)
axis(2, cex.axis=1.1, at=seq(-1,+1, by=.1), las=1)
abline(v=c(1:nlevels(otuCorrs$family)), col="dark gray")
#abline(h=c(.3, -.3), col="dark gray", lty="dashed", lwd=4) #change to something else
abline(h=0, col="black", lwd=2)
abline(h=c(0.25, 0.5, 0.75, -0.25, -0.5, -0.75), col="light gray", lwd=2, lty="dashed")
legend("top", legend=unique(otuCorrs$phyl), pch=c(1, 6, 0, 9, 2, 4), cex=0.7, bty="o", horiz=TRUE)

@@

Figure saved as "corr_avgAbove0.001_topdose2_10x8.pdf".

@@

The code for this is the script "abxD01.barcharts.xOTU.test.r" in the paper's home Github directory. Opening this file, the parameters I changed in the top were:
sortbyphyl<-TRUE
graphbyphyl<-TRUE
abx<-c("5 mg/ml", "0.5 mg/ml", "0.1 mg/ml")
file
fileIDS

The file/fileIDS should be matched by the titration.
Input files on github:
1. abxD01.final.tx.2.subsample.allceftitr.forlogscale.csv
2. abxD01.final.tx.2.subsample.allstreptitr.forlogscale.csv
3. abxD01.final.tx.2.subsample.allvanctitr.forlogscale.csv
4. allceftitr_tx2_barchart_ids.csv
5. allstreptitr_tx2_barchart_ids.csv
6. allvanctitr_tx2_barchart_ids.csv

For this graph I decided to take out the untreated communities to highlight the differences among titration groups. So in r I changed the file a bit:

```{r, eval=FALSE}
file<-file[file$expgroup!="1untrStrep",]
#alternatively for the other titrated antibiotics it's "1untrCef", or "1untrVanc"
```
